---
title: "Balancing Accuracy and Interpretability in Multi-Sensor Fusion through Dynamic Bayesian Networks"
output: html_notebook
---
The complete structure of our BN is composed by:

 - The latent variable 'X' representing the actual position (state) of the obstacle
 
 - The measurable variable 'Si' representing the sensor "i" recording the position (state) of the obstacle. Evidences can be collected. All sensors are listed below in the sensors list.
 
 - 'Mi' define the missingness of the measure of the sensor i.
 
 Sensors list:
 
 - Sr: Range sensors
 
 - Sw_v5: Camera on the wheelchair with yolo v5 model
 
 - Sd_v5: Camera on the drone with yolo v5 model
 
 - Sw_FT: Camera on the wheelchair with yolo v8 fine tuned
 
 - Sd_FT: Camera on the drone with yolo fine tuned
 
All required library to upload
```{r}
#If required, download package for dynamic BN
#install.packages("openxlsx")
#install.packages("bnlearn")
#install.packages("data.table")
#install.packages('dbnR')
#install.packages('visNetwork')
#latest version
#install.packages('devtools')
#devtools::install_github("dkesada/dbnR")
#install.packages("dplyr")
#install.packages("jtools")


library(tidyverse) 
library(grid)
library(graph)
library(Rgraphviz)
library(gRbase)
library(RBGL)
library(bnlearn)
library(magrittr) 
library(dplyr)    
library(plyr)
library(gRain)
library(visNetwork)
library(devtools)
library(dbnR)
library(Metrics)
library(writexl)
library(openxlsx)
library(jtools)
library(stringr)
```

Dataset elaboration for NN and XGBoost to set three timestamp 

```{r}
#import data for range sensor
filter_list<-c('original','rain_1','rain_2','fog_1')
exp_list=c(0,1,2,3,4,5,6,7,8,9,10,11,12,13)
#type_imputation_list=c('imputation_2')
size=3
folder_timeslices3=paste("dataset_timeslices3",sep='')
if (!dir.exists(folder_timeslices3)) dir.create(folder_timeslices3)

for (filter in filter_list){
  for (exp in exp_list){
    Data<- read.csv(paste("imputed_data/imputed_",filter,"_run_",exp,".csv",sep=''))
    Data <- fold_dt(Data, size) 
    #Save new temporized dataset
    #name_dataset=paste("imputed_",filter,"_run_",exp, sep='')
    results_file_name=paste(folder_timeslices3,"/imputed_",filter,"_run_",exp,".csv",sep='')
    write.csv(Data,results_file_name, row.names = TRUE) 
  }
}
```

Function to calculate the AVARAGE model (simple mean of the available outputs of the sensors), and the label of the sensors functioning.
- label_no_sens=1 if no sensor is working 
- label_no_sens=0 otherwise
```{r}
media <- function(dataset,num_exp){
  n <- nrow(dataset)
  media <- vector(,n)
  label_no_sens <- vector(,n)
  for (i in 1:n){
    x <- c()  
    if (dataset$Mw_FT_t_0[i]==0){
      x <- append(x, dataset$Sw_FT_t_0[i])
    }
    if (dataset$Md_FT_t_0[i]==0){
      x <- append(x, dataset$Sd_FT_t_0[i])
    }
    if (dataset$Mw_v5_t_0[i]==0){
      x <- append(x, dataset$Sw_v5_t_0[i])
    }
    if (dataset$Md_v5_t_0[i]==0){
      x <- append(x, dataset$Sd_v5_t_0[i])
    }
    if (dataset$Mr_t_0[i]==0){
      x <- append(x, dataset$Sr_t_0[i])
    }
    if (length(x)==0){

      if(i>1){
        #print('qua')
        media[i] <- media[i-1]
        label_no_sens[i] <- 1
        
      }else{
        print('in media uno step indietro')
        
        if (dataset$Mw_FT_t_1[i]==0){
          x <- append(x, dataset$Sw_FT_t_1[i])
        }
        if (dataset$Md_FT_t_1[i]==0){
          x <- append(x, dataset$Sd_FT_t_1[i])
        }
        if (dataset$Mw_v5_t_1[i]==0){
          x <- append(x, dataset$Sw_v5_t_1[i])
        }
        if (dataset$Md_v5_t_1[i]==0){
          x <- append(x, dataset$Sd_v5_t_1[i])
        }
        if (dataset$Mr_t_1[i]==0){
          x <- append(x, dataset$Sr_t_1[i])
        }
        
        if (length(x)==0){
          print('in media due step indietro')
            if (dataset$Mw_FT_t_2[i]==0){
              x <- append(x, dataset$Sw_FT_t_2[i])
            }
            if (dataset$Md_FT_t_2[i]==0){
              x <- append(x, dataset$Sd_FT_t_2[i])
            }
            if (dataset$Mw_v5_t_2[i]==0){
              x <- append(x, dataset$Sw_v5_t_2[i])
            }
            if (dataset$Md_v5_t_2[i]==0){
              x <- append(x, dataset$Sd_v5_t_2[i])
            }
            if (dataset$Mr_t_2[i]==0){
              x <- append(x, dataset$Sr_t_2[i])
            }
        }
        if (length(x)==0){
          print('ERROR')}
        
        label_no_sens[i] <- 0
        media[i] <- mean(x, na.rm=TRUE)
      }
    }else{
      label_no_sens[i] <- 0
      media[i] <- mean(x, na.rm=TRUE)
    }
  }

  results <- list("label_no_sens" = label_no_sens, "media" = media) 
  return(results)
}
```

Function to create nested folders
```{r}
create_nested_folder <- function(base, ...) {
  path <- file.path(base, ...)
  if (!dir.exists(path)) dir.create(path, recursive = TRUE)
  return(path)
}
```


Set configuration:
```{r}
#type_imputation_list=c('imputation_2')
filter_all<-c('original','rain_1','rain_2','fog_1')
filter_list_val<-filter_all
value_seed=123

# to set the fold for seven fold cross validation:
set_cross=c(4,13,5,12,0,10,3,11,2,9,6,7,1,8)
fold1=c(4,13)
fold2=c(5,12)
fold3=c(0,10)
fold4=c(3,11)
fold5=c(2,9)
fold6=c(6,7)
fold7=c(1,8)
fold_list=list(fold1,fold2,fold3,fold4,fold5,fold6,fold7)
fold_list_name=list('fold1','fold2','fold3','fold4','fold5','fold6','fold7')

#Chose number of temporal slices insert a vector of len n where n is the number of architecture we want to create.
interval_steps=c(3)

#labels of sensors 
list_sensors_label= c('w_FT','d_FT','w_v5','d_v5','r')

#chose if including inter-slices arcs between Si
#inter_arcs_list=c('no_arcs')
#label_acceleration_list=c(TRUE )

#setwd("~/Documents/projects/Bayesian Network/bella copia per condivisione 21_07_2025")  # Example for Windows

```



Upload dataset
```{r}
#for (type_imputation in type_imputation_list){
#  cat("Processing type_imputation:", type_imputation, "\n")
for (filter in filter_all){
  for (exp in set_cross){
      #print(exp)
      
      #url<-paste("~/Documents/projects/Bayesian Network/bella copia per condivisione 21_07_2025/preprocessing_results_02/imputation_2/prepr_data_",filter,"_run_",exp,".csv",sep='')
      
      url <- paste0("imputed_data/imputed_",filter,"_run_",exp,".csv")
      Data<- read.csv(url)
      Data <- Data [c("distance_tracker", "distance_wheelchair_FT", "missing_w_FT","distance_drone_FT", "missing_d_FT","distance_wheelchair_v5", "missing_w_v5","distance_drone_v5", "missing_d_v5","distance_range", "missing_range","all_missing")] %>%
      filter(!is.na(missing_w_FT))  %>%
      filter(!is.na(missing_d_FT))  %>%
      filter(!is.na(missing_w_v5))  %>%
      filter(!is.na(missing_d_v5))  %>%
      filter(!is.na(missing_range))  %>%
      filter(!is.na(distance_tracker))  %>%
      filter(!is.na(distance_wheelchair_FT))  %>%
      filter(!is.na(distance_drone_FT))  %>%
      filter(!is.na(distance_wheelchair_v5))  %>%
      filter(!is.na(distance_drone_v5))  %>%
      filter(!is.na(distance_range))  %>%
      plyr::rename(c("distance_tracker" = "X")) %>%
      plyr::rename(c("distance_wheelchair_FT" = "Sw_FT")) %>%
      plyr::rename(c("missing_w_FT" = "Mw_FT")) %>%
      plyr::rename(c("distance_drone_FT" = "Sd_FT")) %>%
      plyr::rename(c("missing_d_FT" = "Md_FT")) %>%
      plyr::rename(c("distance_wheelchair_v5" = "Sw_v5")) %>%
      plyr::rename(c("missing_w_v5" = "Mw_v5")) %>%
      plyr::rename(c("distance_drone_v5" = "Sd_v5")) %>%
      plyr::rename(c("missing_d_v5" = "Md_v5")) %>%
      plyr::rename(c("distance_range" = "Sr")) %>%
      plyr::rename(c("missing_range" = "Mr")) %>%
      mutate(Mw_FT=factor(Mw_FT,levels=c("0","1")))%>%
      mutate(Md_FT=factor(Md_FT,levels=c("0","1"))) %>%
      mutate(Mw_v5=factor(Mw_v5,levels=c("0","1"))) %>%
      mutate(Md_v5=factor(Md_v5,levels=c("0","1"))) %>%
      mutate(Mr=factor(Mr,levels=c("0","1")))
      
      texto <- paste("Data_",exp, "_",filter,"<- Data", sep='')
      eval(parse(text = texto)) 
  }
}

```


```{r}
CASE='CASE1'

#folder with the results of thee Kalman filter
url_Kalman_filter_pre <- paste0("Kalman_filter_results")
  
# ==== Folders for the results ====
# FOLDER FOR PLOTS
#folder_plot <- create_nested_folder("plots")

# FOLDER FOR PERFORMANCE RESULTS
folder_results <- create_nested_folder("BNresults")

# FOLDER FOR csv plot
folder_csv <- create_nested_folder(folder_results,"data_for_plot/")

# ==== Create dataframe to collect performance ====

rox_names <- unlist(lapply(filter_all, function(f) paste(f, fold_list_name, sep = "_")))
columns_name <- c('KF','baseline')
performance <- data.frame(matrix(NA, nrow = length(rox_names), ncol =length(columns_name)))
rownames(performance) <-  rox_names
colnames(performance) <- columns_name
```

```{r}
#for (inter_arcs in inter_arcs_list){
#  cat("Processing inter_arcs:", inter_arcs, "\n")
#for (label_acceleration in label_acceleration_list){
#  cat("Processing label_acceleration:", label_acceleration, "\n")


for (step in interval_steps){
  cat("Processing step:", step, "\n")
  
  # ==== Create Bayesian network architecture ====
  n_slices <- step
  final_string1=''
  final_string2=''
  #nodes names to be collected in BN plots
  plot_list <- vector("list", length = n_slices)
  
  for (t in 0:(n_slices-1)){
    X_string=paste('X_t_',t,sep='')
    
    if (t==(n_slices-1)){
      # Oldest slice: different structure
      list1=c()
      for (sens in list_sensors_label){
        M_string=paste('M',sens,'_t_',t,sep='')
        S_string=paste('S',sens,'_t_',t,sep='')
        list1= append(list1,S_string,after = length(list1))
        list1= append(list1,M_string,after = length(list1))
        final_string1=paste(final_string1,'[',M_string,']','[',S_string,'|',X_string,':',M_string,']',sep='')
      }
      final_string1=paste('[',X_string,']',final_string1,sep='')
      plot_list[[t+1]] <- list1
    }else if (n_slices!=1){
      # Middle slices (standard structure)
      string2=''
      list2=c()
      
      for (sens in list_sensors_label){
        M_string=paste('M',sens,'_t_',t,sep='')
        S_string=paste('S',sens,'_t_',t,sep='')
        #S_string_old=paste('S',sens,'_t_',t+1,sep='')
        
        list2= append(list2,S_string,after = length(list2))
        list2= append(list2,M_string,after = length(list2))
        
        string2=paste(string2,'[',M_string,']','[',S_string,'|',X_string,':',M_string,']',sep='')
      }
  
      plot_list[[t+1]] <- list2
      string2=paste(string2,'[',X_string,'|X_t_',t+1,sep='')
      if (t!=n_slices-2){
        string2=paste(string2,':X_t_',t+2,sep='')
      }
      final_string2=paste(final_string2,string2,']',sep='')
    }
  }
  # Final string to create BN
  full_network_string=paste(final_string1,final_string2,sep='')
  
  # BN realization
  bn_all=model2network(full_network_string)
  
  # Architecture Plot
  graphviz.plot(bn_all, layout = "dot", shape='circle', fontsize=18, groups = plot_list)

  # ==== CROSS VALIDATION ====
  for (fold in fold_list){
    print(paste("FOLD:", fold))
    train_set<- setdiff(set_cross,as.vector(fold))
    exp_list_validation <- fold
  
    #Creation of the training dataset
      
    size=n_slices
    
    for (exp in train_set){
      Data <- get(paste0("Data_", exp,"_original"))
      Data$all_missing <- NULL
      
      if (size==1){
       names(Data)[names(Data)=='X'] <- 'X_t_0'
      
       for (sens in list_sensors_label){
         texto <-paste0("names(Data)[names(Data)=='S",sens,"'] <- 'S",sens,"_t_0'")
         eval(parse(text = texto)) 
         texto <-paste0("names(Data)[names(Data)=='M",sens,"'] <- 'M",sens,"_t_0'")
         eval(parse(text = texto)) 
       }
       f_dt <- Data 
      }else{
       f_dt <- fold_dt(Data, size) 
      }
      if (exp==train_set[1]){
        f_dt_all_train <- f_dt
      }else if(exp %in% train_set){
        f_dt_all_train <- rbind(f_dt_all_train,f_dt) }
      else{
        print('ERRORE')
        texto <- paste0("f_dt_val_",exp," <- f_dt")
        eval(parse(text = texto)) 
      }
    }
    
    # ==== BN fitting operation with training dataset ====
  
    shuffled_data <- f_dt_all_train[sample(nrow(f_dt_all_train)), ]
    bn_fitted= bn.fit(bn_all, shuffled_data, method = "mle-cg")
    
    
    # ==== BN Validation operation with validation dataset ====
    for (filter in filter_list_val){
      cat("Processing filter:", filter, "\n")  
      size=n_slices
      for (exp in exp_list_validation){
        Data <- get(paste0("Data_", exp,"_",filter))
        if (size==1){
         names(Data)[names(Data)=='X'] <- 'X_t_0'
         names(Data)[names(Data)=='all_missing'] <- 'all_missing_t_0'
         for (sens in list_sensors_label){
           texto <-paste0("names(Data)[names(Data)=='S",sens,"'] <- 'S",sens,"_t_0'")
           eval(parse(text = texto)) 
           texto <-paste0("names(Data)[names(Data)=='M",sens,"'] <- 'M",sens,"_t_0'")
           eval(parse(text = texto)) 
         }
         f_dt <- Data 
        }else{
         f_dt <- fold_dt(Data, size) 
        }
        texto <- paste0("f_dt_val_",exp," <- f_dt")
        eval(parse(text = texto)) 
      }
      
      # ==== Prediction step ====
        
      for (exp in exp_list_validation){
        f_dt_val <- get(paste0("f_dt_val_",exp))
        # Specify columns to remove
        label_missing=f_dt_val$all_missing_t_0
        cols_to_remove <- c(sapply(1:step, function(i) paste('X_t_',i-1,sep='')),sapply(1:step, function(i) paste('all_missing_t_',i-1,sep='')))
        f_dt_val <- f_dt_val %>% select(-all_of(cols_to_remove))
        
        # Prediction operation with cpdist
        mean_preds <- c()
        sd_preds <- c()
        for( q in 1:nrow(f_dt_val)){
          evidence <- as.list(f_dt_val[q,])
          set.seed(value_seed)
          preds <- cpdist(bn_fitted,nodes = "X_t_0",evidence=evidence, method = "lw", query.nodes=names(f_dt_val), debug=FALSE)
          w <-  attr(preds,"weights")
          mean_preds <- append(mean_preds,weighted.mean(preds$X_t_0,w))
          sd_preds <- append(sd_preds,wtd.sd(preds$X_t_0,w))
        }
        texto <-paste0("c11111_exp",exp,"=mean_preds")
        eval(parse(text = texto)) 
        texto <-paste0("c11111_exp",exp,"_uncert=sd_preds")
        eval(parse(text = texto))     
      }
      
      
      # ==== Collect Results and performance metrics ====
    
      for (exp in exp_list_validation){
        #print(paste('exp_list_validation for plots=',exp))
        
        #comparison with Kalman Filter
        url_Kalman_filter <- paste(url_Kalman_filter_pre,"/KF_results_",filter,"_run_",exp,".csv",sep='')
        KF_data<- read.csv(url_Kalman_filter)
         
        KF_predictions <- KF_data[(step:nrow(KF_data)),'predictions']
        KF_uncertainty <- KF_data[(step:nrow(KF_data)),'uncertainty']
        
        df <- get(paste0("f_dt_val_",exp))
        c11111 <- get(paste0("c11111_exp",exp))
        c11111_sd <- get(paste0("c11111_exp",exp,"_uncert"))
        baseline<- media(df,exp)
        label_sensors<- baseline$label_no_sens
        baseline <- baseline$media
        index=1:nrow(df)
      
        
        # Results collection for each validation dataset/run
      
        
        if (exp==exp_list_validation[1]){
          c11111_last <- c11111
          c11111_sd_last <- c11111_sd
          baseline_last <- baseline
          df_last <- df
          KF_predictions_last <- KF_predictions
        }else{
          c11111_final <- c(c11111_last,c11111)
          c11111_sd_final <- c(c11111_sd_last,c11111_sd)
          baseline_final <- c(baseline_last,baseline)
          KF_predictions_final <- c(KF_predictions_last,KF_predictions)
          df_final <- rbind(df_last,df)
        }
        
        data_for_plots=data.frame(1:length(c11111),
                                  baseline,
                                  c11111,c11111_sd,
                                  df$X_t_0,
                                  df$Sw_FT_t_0*(1-(as.numeric(df$Mw_FT_t_0)-1)),
                                  df$Sd_FT_t_0*(1-(as.numeric(df$Md_FT_t_0)-1)),
                                  df$Sw_v5_t_0*(1-(as.numeric(df$Mw_v5_t_0)-1)),
                                  df$Sd_v5_t_0*(1-(as.numeric(df$Md_v5_t_0)-1)),
                                  df$Sr_t_0*(1-(as.numeric(df$Mr_t_0)-1)),
                                  KF_predictions,KF_uncertainty) 
        
        colnames(data_for_plots) <- c('time','average','BN_mean','BN_sd','X_t_0','Sw_FT_t_0','Sd_FT_t_0','Sw_v5_t_0','Sd_v5_t_0','Sr_t_0','KF_predictions','KF_sd')
        data_for_plots[data_for_plots==0] <- NA
        
        # ==== Save Results====
        write.csv(data_for_plots, paste0(folder_csv,"data_for_plot_",filter,"_exp",exp,".csv"),row.names = FALSE)
      }
    
      num_folder<- which(sapply(fold_list, function(x) identical(x, fold)))
      
      performance[paste0(filter,'_fold',num_folder),paste0('step_',step)] <-  rmse(df_final$X_t_0,c11111_final)
      performance[paste0(filter,'_fold',num_folder),paste0('baseline')] <- rmse(df_final$X_t_0,baseline_final)
      performance[paste0(filter,'_fold',num_folder),paste0('KF')] <- rmse(df_final$X_t_0,KF_predictions_final)
    }
    print(paste0('FINE fold',fold))
  }
  print(paste0('FINE step',step))
}

# ==== Save Metrics====
write.csv(performance, paste0(folder_results,"/RMSE_comparison_results.csv"))
```


```{r}
#write.xlsx(performance, paste0(folder_results,"/results_crossvalFINALE.xlsx"), rowNames = TRUE)
#write.xlsx(performance_arcs, paste0(folder_results,"/results_crossvalFINALE_with arcs.xlsx"), rowNames = TRUE)
#write.xlsx(performance_noarcs, paste0(folder_results,"/results_crossvalFINALE_withoutarcs.xlsx"), rowNames = TRUE)

```

